### 2017-11-24 templates/livecoms-template-lessonslearned.tex
Add templates for all other document types, as well as a sample document. Update READMEs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
%%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE
\documentclass[9pt,lessons]{livecoms}
% Use the 'onehalfspacing' option for 1.5 line spacing
% Use the 'doublespacing' option for 2.0 line spacing
% Use the 'lineno' option for adding line numbers.
% The 'bestpractices' option for indicates that this is a best practices guide.
% Omit the bestpractices option to remove the marking as a LiveCoMS paper.
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\graphicspath{{figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% IMPORTANT USER CONFIGURATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\versionnumber}{1.3}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/myaccount/homegithubrepository}}  %this should be the main github repository for this article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{This is the title : v\versionnumber}

\author[1*]{Firstname Middlename Surname}
\author[1,2\authfn{1}\authfn{3}]{Firstname Middlename Familyname}
\author[2\authfn{1}\authfn{4}]{Firstname Initials Surname}
\author[2*]{Firstname Surname}
\affil[1]{Institution 1}
\affil[2]{Institution 2}

\corr{email1@example.com}{FMS}  % Correspondence emails.  FMS and FS are the appropriate authors initials.
\corr{email2@example.com}{FS}

\contrib[\authfn{1}]{These authors contributed equally to this work}
\contrib[\authfn{2}]{These authors also contributed equally to this work}

\presentadd[\authfn{3}]{Department, Institute, Country}
\presentadd[\authfn{4}]{Department, Institute, Country}

\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}
\maketitle

\begin{abstract}
This particular document provides a skeleton illustrating key sections for a Lessons Learned document.
Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
Here we also assume familiarity with LaTeX and knowledge of how to include figures, tables, etc.; if you want examples, see the sample just referenced.

In your work, in this particular slot, please provide an abstract of no more than 250 words.
Your abstract should explain the main contributions of your article, and should not contain any material that is not included in the main text.
Please note that your abstract, plus the authorship material following it, must not extend beyond the title page or modifications to the LaTeX class will likely be needed.
\end{abstract}

\end{frontmatter}

\section{Introduction}

Here you would explain what problem you are tackling and briefly motivate your work.
Since a Lessons Learned document provides negative results in a particular category, you will want to clearly define your category and also motivate the application of your work to your particular problem.

We suggest that you envision your document potentially extending in the future to include other Lessons Learned in the same category, so you may want to organize your work in a way which would allow this type of extension (e.g. providing a general introduction then using subsections to explain your particular problem) -- for example, describing potential of mean force calculations generally, then in a subsection motivating your particular application of them.

In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.

A Lessons Learned document is a regular scientific article, and should be prepared with similar organization.
Sample typical section headers follow.

\section{Theory and Methods}

You should use a section (or separate sections, if this yields better organization) to describe the theory and methods relevant to your work.

This should include all information required for a reader to reproduce the results, including software version numbers and links to any in-house files or scripts

\section{Results}

You would give your results here.
Since the goal here is not just to present negative results, but also to provide lessons learned from doing so, you will want to make sure you have data to show why your approach failed.

\section{Discussion}

Discussion should highlight the lessons learned and give tips for avoiding future mistakes.
Try to tell readers what you wish you had known before starting.

\section{Conclusions}

Give conclusions.

\section{Author Contributions}
%%%%%%%%%%%%%%%%
% This section mustt describe the actual contributions of
% author. Since this is an electronic-only journal, there is
% no length limit when you describe the authors' contributions,
% so we recommend describing what they actually did rather than
% simply categorizing them in a small number of
% predefined roles as might be done in other journals.
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
% for more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%%

(Explain the contributions of the different authors here)

% We suggest you preserve this comment:
For a more detailed description of author contributions,
see the GitHub issue tracking and changelog at \githubrepository.

\section{Other Contributions}
%%%%%%%%%%%%%%%
% You should include all people who have filed issues that were
% accepted into the paper, or that upon discussion altered what was in the paper.
% Multiple significant contributions might mean that the contributor
% should be moved to authorship at the discretion of the a
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io for
% more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%

(Explain the contributions of any non-author contributors here)
% We suggest you preserve this comment:
For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.

\section{Potentially Conflicting Interests}
%%%%%%%
%Declare any potentially competing interests, financial or otherwise
%%%%%%%

Declare any potentially conflicting interests here, whether or not they pose an actual conflict in your view.

\section{Funding Information}
%%%%%%%
% Authors should acknowledge funding sources here. Reference specific grants.
%%%%%%%
FMS acknowledges the support of NSF grant CHE-1111111.

\bibliography{livecoms-sample}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\appendix

\end{document}

### 2017-11-24 templates/livecoms-template-reviews.tex
Add templates for all other document types, as well as a sample document. Update READMEs.

--- 2017-11-24 templates/livecoms-template-lessonslearned.tex
@@
 %%% PREAMBLE
-\documentclass[9pt,lessons]{livecoms}
+\documentclass[9pt,review]{livecoms}
 % Use the 'onehalfspacing' option for 1.5 line spacing
@@
 \begin{abstract}
-This particular document provides a skeleton illustrating key sections for a Lessons Learned document.
+This particular document provides a skeleton illustrating key sections for a Perpetual Review document.
 Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
@@
 
-Here you would explain what problem you are tackling and briefly motivate your work.
-Since a Lessons Learned document provides negative results in a particular category, you will want to clearly define your category and also motivate the application of your work to your particular problem.
+Here you would explain what area you are reviewing, and motivate your review.
 
-We suggest that you envision your document potentially extending in the future to include other Lessons Learned in the same category, so you may want to organize your work in a way which would allow this type of extension (e.g.
-providing a general introduction then using subsections to explain your particular problem) -- for example, describing potential of mean force calculations generally, then in a subsection motivating your particular application of them.
+In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify.
 
-In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.
+\section{Review organization}
 
-A Lessons Learned document is a regular scientific article, and should be prepared with similar organization.
-Sample typical section headers follow.
+In general the organization of your review is up to you.
+LiveCoMS reviews are similar to traditional reviews except that they are updatable.
 
-\section{Theory and Methods}
+\subsection{Perpetual review}
 
-You should use a section (or separate sections, if this yields better organization) to describe the theory and methods relevant to your work.
+You will probably want to devote at least a small amount of space explaining to your readers what a Perpetual Review is and highlighting how they can get involved via your git issue tracker or by communicating with you directly.
+In contrast to a traditional review, here it \emph{will} help you if the community points out other, related work that perhaps should be cited, since it reduces the need for you to dig deeply into the literature to keep your work up to date.
 
-This should include all information required for a reader to reproduce the results, including software version numbers and links to any in-house files or scripts
+Tutorials should endeavor to cover the specific task at hand, and also highlight how the steps might need to be modified (or additional care might need to be taken at particular points) to handle more general cases.
 
-\section{Results}
+The scope of the tutorial, as well as the expected proficiencies / outcomes for researchers who complete the tutorial, should be clearly defined.
+This will often happen in a specific section or subsection in the article itself.
 
-You would give your results here.
-Since the goal here is not just to present negative results, but also to provide lessons learned from doing so, you will want to make sure you have data to show why your approach failed.
+\section{Prerequisites}
 
-\section{Discussion}
-
-Discussion should highlight the lessons learned and give tips for avoiding future mistakes.
-Try to tell readers what you wish you had known before starting.
+In keeping with other types of articles at LiveCoMS, it may be useful (at your discretion) to note any prerequisites your article assumes and who it is targeted at.
 
@@
 
-Give conclusions.
+It's appropriate to wrap up with discussion and conclusions.
 

### 2017-11-26 templates/livecoms-template-computationalcomparisons.tex

--- 2017-11-24 templates/livecoms-template-lessonslearned.tex
@@
 %%% PREAMBLE
-\documentclass[9pt,lessons]{livecoms}
+\documentclass[9pt,comparison]{livecoms}
 % Use the 'onehalfspacing' option for 1.5 line spacing
@@
 \begin{abstract}
-This particular document provides a skeleton illustrating key sections for a Lessons Learned document.
+This particular document provides a skeleton illustrating key sections for a Computational Comparisons document.
 Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
@@
 
-Here you would explain what problem you are tackling and briefly motivate your work.
-Since a Lessons Learned document provides negative results in a particular category, you will want to clearly define your category and also motivate the application of your work to your particular problem.
+Here you would explain what problem you are tackling and briefly motivate your work; why is a comparison needed here?
+What do the tools you are comparing do?
 
-We suggest that you envision your document potentially extending in the future to include other Lessons Learned in the same category, so you may want to organize your work in a way which would allow this type of extension (e.g.
-providing a general introduction then using subsections to explain your particular problem) -- for example, describing potential of mean force calculations generally, then in a subsection motivating your particular application of them.
+In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify but we retain an example on algorithms and pseudocode.
 
-In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.
+\section{Methodology}
 
-A Lessons Learned document is a regular scientific article, and should be prepared with similar organization.
-Sample typical section headers follow.
+Here (and in other sections if needed) you would describe how you compared the software in question, with subsections as needed.
+Perhaps you may also need to comment here on why you did not include certain pieces of software or how these might be included in the future.
 
-\section{Theory and Methods}
+\subsection{Physical systems and properties}
 
-You should use a section (or separate sections, if this yields better organization) to describe the theory and methods relevant to your work.
+What did you calculate, on what systems? Why did you make that choice?
 
-This should include all information required for a reader to reproduce the results, including software version numbers and links to any in-house files or scripts
+\subsection{Methodology for comparison}
+How did you compute what you're comparing? 
+What software versions did you use?
+How would the software be accessed?
+
+\subsubsection{Method selection}
+How did you select which protocols to use? How are you certain these are representative? Would a naive user get similar performance?
+
+\subsubsection{Supporting files}
+Are there run scripts or files others would need to reproduce your work? Are these provided in your GitHub repo, or how would they be accessed? 
+How exactly would someone reproduce/extend this work?
+
+\subsubsection{Error analysis}
+You may wish to devote particular attention to error analysis.
 
@@
 
-You would give your results here.
-Since the goal here is not just to present negative results, but also to provide lessons learned from doing so, you will want to make sure you have data to show why your approach failed.
+You would give results.
 
-\section{Discussion}
+\section{Discussion and conclusions}
 
-Discussion should highlight the lessons learned and give tips for avoiding future mistakes.
-Try to tell readers what you wish you had known before starting.
-
-\section{Conclusions}
-
-Give conclusions.
+You may perhaps wish to revisit the issue of what performance a naive user might expect, here.
 

### 2018-06-15 templates/livecoms-template-bestpractices.tex

--- 2017-11-24 templates/livecoms-template-lessonslearned.tex
@@
 %%% PREAMBLE
-\documentclass[9pt,lessons]{livecoms}
+\documentclass[9pt,bestpractices]{livecoms}
 % Use the 'onehalfspacing' option for 1.5 line spacing
@@
 % Omit the bestpractices option to remove the marking as a LiveCoMS paper.
+% Include "pubversion" option when the volume/article/DOI information is known
 % Please note that these options may affect formatting.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% PUBLICATION INFORMATION
+%%% Fill out these parameters when available
+%%% These are used when the "pubversion" option is invoked
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pubDOI{10.XXXX/A34N348}
+\pubvolume{23}
+\pubyear{2018}
+\articlenum{A34N348}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%% ARTICLE START
@@
 \begin{abstract}
-This particular document provides a skeleton illustrating key sections for a Lessons Learned document.
+This particular document provides a skeleton illustrating key sections for a Best Practices document.
 Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
@@
 Here you would explain what problem you are tackling and briefly motivate your work.
-Since a Lessons Learned document provides negative results in a particular category, you will want to clearly define your category and also motivate the application of your work to your particular problem.
-
-We suggest that you envision your document potentially extending in the future to include other Lessons Learned in the same category, so you may want to organize your work in a way which would allow this type of extension (e.g.
-providing a general introduction then using subsections to explain your particular problem) -- for example, describing potential of mean force calculations generally, then in a subsection motivating your particular application of them.
 
@@
 
-A Lessons Learned document is a regular scientific article, and should be prepared with similar organization.
-Sample typical section headers follow.
+\section{Prerequisites}
 
-\section{Theory and Methods}
+Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
+Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.
 
-You should use a section (or separate sections, if this yields better organization) to describe the theory and methods relevant to your work.
+\section{Checklist}
+Here we use a full-page checklist with multiple sections, so it will appear on a separate page of the sample PDF.
+Other checklist formats are possible, as shown in the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
 
-This should include all information required for a reader to reproduce the results, including software version numbers and links to any in-house files or scripts
+Your checklist should include a succinct list of steps that people should follow when carrying out the task in question.
+This is provided to ensure certain basic standards are followed and common but critical major errors are avoided.
+Note that a checklist is not intended to cover \emph{all} important steps, but rather focus on the most common reasons for failure or incorrect results, or issues which are particularly crucial.
 
-\section{Results}
+% This provides a checklist which
+% - spans a full page
+% - consists of multiple sub-checklists
+% - exists on a separate page
+% This style of checklist will be especially helpful if you want to encourage readers to print and use your checklist in practice, as they
+% can easily print it without also printing other material from your manuscript. However, other styles of checklist are also possible (below).
+\begin{Checklists*}[p!]
 
-You would give your results here.
-Since the goal here is not just to present negative results, but also to provide lessons learned from doing so, you will want to make sure you have data to show why your approach failed.
+\begin{checklist}{First list}
+\textbf{You can easily make full width checklists that take a whole page}
+\begin{itemize}
+\item Items in your checklist can and should reference sections where the cited issues are discussed in detail, such as Section~\ref{sec:reference_this}
+\item Also remember
+\item And finally
+\end{itemize}
+\end{checklist}
 
-\section{Discussion}
+\begin{checklist}{Second list}
+\textbf{This is some further description.}
+\begin{itemize}
+\item First thing
+\item Also remember
+\item And finally
+\end{itemize}
+\end{checklist}
 
-Discussion should highlight the lessons learned and give tips for avoiding future mistakes.
-Try to tell readers what you wish you had known before starting.
+\begin{checklist}{Third list}
+\textbf{This is some further description.}
+\begin{itemize}
+\item First thing
+\item Also remember
+\item And finally
+\end{itemize}
+\end{checklist}
 
-\section{Conclusions}
+\begin{checklist}{Fourth list}
+\textbf{This is some further description.}
+\begin{itemize}
+\item First thing
+\item Also remember
+\item And finally
+\end{itemize}
+\end{checklist}
 
-Give conclusions.
+\end{Checklists*}
+
+\section{Rationale}
+
+Your Rationale section, or sections, can follow or precede your checklist (we expect that often, following the checklist will be preferable) and provide the necessary rationale for the checklist, and act as more complete \emph{best practices} description.
+This should include 1) significant detail as to the possible alternative ways to accomplish a given task, 2) description of advantages and disadvantages of the various approaches, and 3) significant literature documentation about reasons for choices.
+
+\subsection{Algorithms and Pseudocode}
+\label{sec:reference_this}
+
+The \texttt{algpseudocode} and \texttt{algorithms} packages is loaded by the document class.
+\ALG{euclid} was taken directly from the package documentation.
+(Please do not load \texttt{algorithm2e}; it's not compatible with \texttt{algpseudocde} nor \texttt{algorithms}!)
+
+\begin{algorithm}
+\caption{Euclid's algorithm}\label{alg:euclid}
+\begin{algorithmic}%[1]  %% uncomment to enable line numbers
+\Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
+   \State $r\gets a\bmod b$
+   \While{$r\not=0$}\Comment{We have the answer if r is 0}
+      \State $a\gets b$
+      \State $b\gets r$
+      \State $r\gets a\bmod b$
+   \EndWhile\label{euclidendwhile}
+   \State \textbf{return} $b$\Comment{The gcd is b}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 

### 2018-06-15 templates/livecoms-template-tutorials.tex

--- 2017-11-24 templates/livecoms-template-lessonslearned.tex
@@
 %%% PREAMBLE
-\documentclass[9pt,lessons]{livecoms}
+\documentclass[9pt,tutorial]{livecoms}
 % Use the 'onehalfspacing' option for 1.5 line spacing
@@
 \begin{abstract}
-This particular document provides a skeleton illustrating key sections for a Lessons Learned document.
+This particular document provides a skeleton illustrating key sections for a Tutorial document.
 Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
@@
 
-A Lessons Learned document is a regular scientific article, and should be prepared with similar organization.
-Sample typical section headers follow.
+\subsection{Scope}
 
-\section{Theory and Methods}
+Tutorials should endeavor to cover the specific task at hand, and also highlight how the steps might need to be modified (or additional care might need to be taken at particular points) to handle more general cases.
 
-You should use a section (or separate sections, if this yields better organization) to describe the theory and methods relevant to your work.
+The scope of the tutorial, as well as the expected proficiencies / outcomes for researchers who complete the tutorial, should be clearly defined.
+This will often happen in a specific section or subsection in the article itself.
 
-This should include all information required for a reader to reproduce the results, including software version numbers and links to any in-house files or scripts
+\section{Prerequisites}
 
-\section{Results}
+Here you would identify prerequisites/background knowledge that are assumed by your work, as well as any software/license requirements.
 
-You would give your results here.
-Since the goal here is not just to present negative results, but also to provide lessons learned from doing so, you will want to make sure you have data to show why your approach failed.
+\subsection{Background knowledge}
+Tutorials should clearly define what concepts or abilities researchers will need to complete the tutorial (e.g., some proficiency in Python; experience with Jupyter notebooks; knowledge of classical MD; etc).
 
-\section{Discussion}
+\subsection{Software/system requirements}
+Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.).
+Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
 
-Discussion should highlight the lessons learned and give tips for avoiding future mistakes.
-Try to tell readers what you wish you had known before starting.
+\section{Content and links}
 
-\section{Conclusions}
+A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
+You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.
 
-Give conclusions.
+\section{Checklists}
+Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
+Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
+One example is shown here.
+
+% Here is a single-column checklist that consists of multiple sub-checklists
+\begin{Checklists}
+
+\begin{checklist}{A list}
+\textbf{Single-column checklists are also straightforward by removing the asterisk}
+\begin{itemize}
+\item First thing let's do an item which breaks across lines to see how that looks
+\item Also remember
+\item And finally
+\end{itemize}
+\end{checklist}
+
+\begin{checklist}{Another list}
+\textbf{This is some further description.}
+\begin{itemize}
+\item First thing
+\item Also remember
+\item And finally
+\end{itemize}
+\end{checklist}
+
+\end{Checklists}
 

### 2019-10-31 templates/livecoms-template-bestpractices.tex
Changed version format in title per issue tracker
Changed date in publication metadata to international 'longform' style
Added ASAPversion class option, updated templates, and recompiled all templates
Add info to README/tempaltes on float specifiers and representative images

@@
 % Use the 'lineno' option for adding line numbers.
+% Use the "ASAPversion' option following article acceptance to add the DOI and relevant dates to the document footer.
+% Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
 % The 'bestpractices' option for indicates that this is a best practices guide.
 % Omit the bestpractices option to remove the marking as a LiveCoMS paper.
-% Include "pubversion" option when the volume/article/DOI information is known
 % Please note that these options may affect formatting.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{This is the title : v\versionnumber}
+\title{This is the title [Article v\versionnumber]}
 
@@
 \corr{email2@example.com}{FS}
+
+\orcid{Author 1 name}{AAAA-BBBB-CCCC-DDDD}
+\orcid{Author 2 name}{EEEE-FFFF-GGGG-HHHH}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\pubDOI{10.XXXX/A34N348}
-\pubvolume{23}
-\pubyear{2018}
-\articlenum{A34N348}
+\pubDOI{10.XXXX/YYYYYYY}
+\pubvolume{<volume>}
+\pubissue{<issue>}
+\pubyear{<year>}
+\articlenum{<number>}
+\datereceived{Day Month Year}
+\dateaccepted{Day Month Year}
 
@@
 In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.
+
+Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications.
+Usually, this would be one of your figures, but it must also be uploaded separately upon article submission.
+We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).
+
+Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so avoiding specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
 
@@
 
+\section*{Author Information}
+\makeorcid
+
 \bibliography{livecoms-sample}

### 2019-10-31 templates/livecoms-template-lessonslearned.tex
Added class option to include journal publication information in article footer. Sample documents updated to include parameter definitions and instructions regarding pubversion class option.
Changed version format in title per issue tracker
Added ORCID macro (simple version), updated TeX templates, and recompiled PDFs
Added ASAPversion class option, updated templates, and recompiled all templates
Add info to README/tempaltes on float specifiers and representative images

@@
 % Use the 'lineno' option for adding line numbers.
+% Use the "ASAPversion' option following article acceptance to add the DOI and relevant dates to the document footer.
+% Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
 % The 'bestpractices' option for indicates that this is a best practices guide.
@@
 
+\orcid{Author 1 name}{AAAA-BBBB-CCCC-DDDD}
+\orcid{Author 2 name}{EEEE-FFFF-GGGG-HHHH}
+
 \contrib[\authfn{1}]{These authors contributed equally to this work}
@@
 \blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% PUBLICATION INFORMATION
+%%% Fill out these parameters when available
+%%% These are used when the "pubversion" option is invoked
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pubDOI{10.XXXX/YYYYYYY}
+\pubvolume{<volume>}
+\pubissue{<issue>}
+\pubyear{<year>}
+\articlenum{<number>}
+\datereceived{Day Month Year}
+\dateaccepted{Day Month Year}
 
@@
 Sample typical section headers follow.
+
+Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications.
+Usually, this would be one of your figures, but it must also be uploaded separately upon article submission.
+We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).
+
+Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so avoiding specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
 

### 2019-10-31 templates/livecoms-template-reviews.tex
Modified class file to put publication information in footer; updated sample TeX files to include necessary definitions; recompiled sample-document
Added issue number to article footer and updated sample TeX files
Added ORCID macro (simple version), updated TeX templates, and recompiled PDFs
Added ASAPversion class option, updated templates, and recompiled all templates
Add info to README/tempaltes on float specifiers and representative images

@@
 In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify.
+
+Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications.
+Usually, this would be one of your figures, but it must also be uploaded separately upon article submission.
+We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).
+
+Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so avoiding specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
 

### 2019-10-31 templates/livecoms-template-softwareanalyses.tex
renaming section.
Add info to README/tempaltes on float specifiers and representative images

--- 2017-11-26 templates/livecoms-template-computationalcomparisons.tex
@@
 %%% PREAMBLE
-\documentclass[9pt,comparison]{livecoms}
+\documentclass[9pt,software]{livecoms}
 % Use the 'onehalfspacing' option for 1.5 line spacing
@@
 In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify but we retain an example on algorithms and pseudocode.
+
+Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications.
+Usually, this would be one of your figures, but it must also be uploaded separately upon article submission.
+We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).
+
+Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so avoiding specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
 

### 2019-11-04 templates/livecoms-template-bestpractices.tex
Mild grammatical changes

@@
 
-Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so avoiding specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
+Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so please avoid specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
 

### 2021-03-17 templates/livecoms-template-training.tex

--- 2019-11-04 templates/livecoms-template-tutorials.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
+%%% LIVECOMS ARTICLE TEMPLATE FOR TRAINING ARTICLE
 %%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
@@
 %%% PREAMBLE
-\documentclass[9pt,tutorial]{livecoms}
+\documentclass[9pt,training]{livecoms}
 % Use the 'onehalfspacing' option for 1.5 line spacing
@@
 % Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
-% The 'bestpractices' option for indicates that this is a best practices guide.
+% The 'training' option for indicates that this is a training article.
 % Omit the bestpractices option to remove the marking as a LiveCoMS paper.
@@
 \begin{abstract}
-This particular document provides a skeleton illustrating key sections for a Tutorial document.
+This particular document provides a skeleton illustrating key sections for a Training article.
 Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
@@
 
-Tutorials should endeavor to cover the specific task at hand, and also highlight how the steps might need to be modified (or additional care might need to be taken at particular points) to handle more general cases.
+Training articles are largely self-contained and introduce readers to the theory and/or typical procedures for a particular aspect of molecular simulation.
 
-The scope of the tutorial, as well as the expected proficiencies / outcomes for researchers who complete the tutorial, should be clearly defined.
+Training articles need not refer to specific software commands and procedures but may do so to further the pedagogical goals of the paper.
+
+As needed, training articles are likely to display more details of calculations than typically are shown in research articles, for example, noting any mathematical principles or “tricks” used to derive key results.
+
+The scope of the training article should be clearly defined.
 This will often happen in a specific section or subsection in the article itself.
@@
 
-Here you would identify prerequisites/background knowledge that are assumed by your work, as well as any software/license requirements.
+Training articles should clearly state the target audience and knowledge prerequisites.
+Key prerequisites should be noted in the article abstract to permit readers to rapidly ascertain an articles suitability.
 
 \subsection{Background knowledge}
-Tutorials should clearly define what concepts or abilities researchers will need to complete the tutorial (e.g., some proficiency in Python; experience with Jupyter notebooks; knowledge of classical MD; etc).
+
+Although the authors may imagine a particular career-level (e.g., undergraduate or graduate), given the diversity of disciplinary curricula, it is more important to specify precisely any knowledge prerequisites (e.g., vector calculus, basic thermodynamics).
 
 \subsection{Software/system requirements}
-Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.).
-Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
+
+If a particular software or programming environment plays a central role in the article, that should be specified.
 
@@
 
-A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
+A training article may on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
 You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.
@@
 \section{Checklists}
-Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
+Training articles do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
 Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.

### 2021-06-09 templates/paper.tex
filled in some of the basic info, and wrote an abstract

--- 2021-03-17 templates/livecoms-template-training.tex
@@
 
-\newcommand{\versionnumber}{1.3}  % you should update the minor version number in preprints and major version number of submissions.
-\newcommand{\githubrepository}{\url{https://github.com/myaccount/homegithubrepository}}  %this should be the main github repository for this article
+\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
+\newcommand{\githubrepository}{\url{https://github.com/GrossfieldLab/article_templates}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{This is the title [Article v\versionnumber]}
+\title{How to be a good member of a scientific software community [Article v\versionnumber]}
 
-\author[1*]{Firstname Middlename Surname}
-\author[1,2\authfn{1}\authfn{3}]{Firstname Middlename Familyname}
-\author[2\authfn{1}\authfn{4}]{Firstname Initials Surname}
-\author[2*]{Firstname Surname}
-\affil[1]{Institution 1}
-\affil[2]{Institution 2}
+\author[1*]{Alan Grossfield}
+%\author[1,2\authfn{1}\authfn{3}]{Firstname Middlename Familyname}
+%\author[2\authfn{1}\authfn{4}]{Firstname Initials Surname}
+%\author[2*]{Firstname Surname}
+\affil[1]{University of Rochester Medical Center, Department of Biochemistry and Biophysics}
+%\affil[2]{Institution 2}
 
-\corr{email1@example.com}{FMS}  % Correspondence emails.  FMS and FS are the appropriate authors initials.
-\corr{email2@example.com}{FS}
+\corr{alan_grossfield@urmc.rochester.edu}{AG}  % Correspondence emails.  FMS and FS are the appropriate authors initials.
+%\corr{email2@example.com}{FS}
 
-\orcid{Author 1 name}{AAAA-BBBB-CCCC-DDDD}
-\orcid{Author 2 name}{EEEE-FFFF-GGGG-HHHH}
+\orcid{Alan Grossfield}{0000-0002-5877-2789}
+%\orcid{Author 2 name}{EEEE-FFFF-GGGG-HHHH}
 
-\contrib[\authfn{1}]{These authors contributed equally to this work}
-\contrib[\authfn{2}]{These authors also contributed equally to this work}
+%\contrib[\authfn{1}]{These authors contributed equally to this work}
+%\contrib[\authfn{2}]{These authors also contributed equally to this work}
 
-\presentadd[\authfn{3}]{Department, Institute, Country}
-\presentadd[\authfn{4}]{Department, Institute, Country}
+%\presentadd[\authfn{3}]{Department, Institute, Country}
+%\presentadd[\authfn{4}]{Department, Institute, Country}
 
@@
 \begin{abstract}
-This particular document provides a skeleton illustrating key sections for a Training article.
-Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
-Here we also assume familiarity with LaTeX and knowledge of how to include figures, tables, etc.; if you want examples, see the sample just referenced.
 
-In your work, in this particular slot, please provide an abstract of no more than 250 words.
-Your abstract should explain the main contributions of your article, and should not contain any material that is not included in the main text.
-Please note that your abstract, plus the authorship material following it, must not extend beyond the title page or modifications to the LaTeX class will likely be needed.
+Software is ubiquitous in modern science --- almost any project, in almost any
+discipline, requires some code to work. However, many (or even most) scientists
+are not programmers, and must rely on programs written and maintained by others.
+As a result, a crucial but often neglected part of a scientist's training is
+learning how to use new tools, and how to exist as part of a community of users.
+This article will discuss key behaviors that can make the experience quicker,
+more efficient, and more pleasant for the user and developer alike.
+
 \end{abstract}

### 2021-06-09 templates/paper.tex
added my standard bibliography file
wrote checklists

@@
 
-\begin{checklist}{A list}
-\textbf{Single-column checklists are also straightforward by removing the asterisk}
+\begin{checklist}{Good community member}
 \begin{itemize}
-\item First thing let's do an item which breaks across lines to see how that looks
-\item Also remember
-\item And finally
+\item Tries to solve problem themselves first
+\item Asks for help in the right place
+\item Writes informative bug reports
+\item Cites and acknowledges software appropriately
+\item Contributes to the community
+\item Treats fellow members and developers with courtesy and respect
 \end{itemize}
@@
 
-\begin{checklist}{Another list}
-\textbf{This is some further description.}
+\begin{checklist}{Poor community member}
 \begin{itemize}
-\item First thing
-\item Also remember
-\item And finally
+\item Doesn't read the manual or search the internet before asking for help
+\item Doesn't use the correct venue to ask for help
+\item Writes vague or unhelpful bug reports, or doesn't respond to questions
+\item Is rude or demanding when requesting support
+\item Treats fellow community members disrespectfully
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%
-
-(Explain the contributions of the different authors here)
+The initial version of this paper was written by Alan Grossfield.
 
@@
 
-(Explain the contributions of any non-author contributors here)
 % We suggest you preserve this comment:
@@
 
-Declare any potentially conflicting interests here, whether or not they pose an actual conflict in your view.
+Alan Grossfield serves as a consultant to two companies, Moderna Therapeutics
+and Atelerix Life Sciences.
 
@@
 %%%%%%%
-FMS acknowledges the support of NSF grant CHE-1111111.
+This work supported in part by NIH R21GM138970 to AG.
 
@@
 
-\bibliography{livecoms-sample}
+\bibliography{bibliography}
 

### 2021-06-09 templates/paper.tex
removed some boilerplate

@@
 \section{Checklists}
-Training articles do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
-Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
-One example is shown here.
-
 % Here is a single-column checklist that consists of multiple sub-checklists

### 2021-06-11 templates/paper.tex
removed boilerplate, added a third checklist

@@
 
-Here you would explain what problem you are tackling and briefly motivate your work.
-
-In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.
-
-Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications.
-Usually, this would be one of your figures, but it must also be uploaded separately upon article submission.
-We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).
-
-Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so please avoid specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
-
-\subsection{Scope}
-
-Training articles are largely self-contained and introduce readers to the theory and/or typical procedures for a particular aspect of molecular simulation.
-
-Training articles need not refer to specific software commands and procedures but may do so to further the pedagogical goals of the paper.
-
-As needed, training articles are likely to display more details of calculations than typically are shown in research articles, for example, noting any mathematical principles or “tricks” used to derive key results.
-
-The scope of the training article should be clearly defined.
-This will often happen in a specific section or subsection in the article itself.
+reveal the hidden curriculum -- what's expected of a software consumer, how to
+ask for help, how to contribute
 
@@
 \item Treats fellow community members disrespectfully
+\end{itemize}
+\end{checklist}
+
+\begin{checklist}{A good community}
+\begin{itemize}
+\item Helps users solve their problems
+\item Is friendly and supportive when responding to questions
+\item Encourages participation from users of all experience levels
+\item Encourages respectful treatment of all community members, and calls out disrespectful behavior
 \end{itemize}

### 2021-06-11 templates/paper.tex
wrote a couple of paragraphs of introduction

@@
 
-reveal the hidden curriculum -- what's expected of a software consumer, how to
-ask for help, how to contribute
+Most practicing scientists (even the ones who write code as part of their
+research) spend more time as consumers as opposed to producers of software.
+Moreover, we are constantly learning new skills and solving new problems, which
+often means learning to use new programs or new aspects of large packages. This,
+combined with the complex nature of scientific software (and the often low
+quality of the associated documentation) means we have to ask for help.
+Sometimes it's because we don't know how to accomplish a specific task. Others,
+it's because there's a feature we need that isn't implemented. Inevitably, there
+are bugs.
+
+In all of these cases, it is necessary to interact with the folks who develop
+and support the code. How you go about it has an enormous impact on your
+likelihood of success and how you are viewed, but this is not something that is
+explicitly taught in most places.  {\emph The goal of this work is to reveal the hidden
+curriculum -- what's expected of a software consumer, how to ask for help, how
+to contribute productively to a software community (regardless of whether you
+can write code).}
+
+There lots of articles on best practices in software development, including open
+source.
 
@@
 \item Is friendly and supportive when responding to questions
+\item Is receptive to suggestions and critiques, regardless of the source
 \item Encourages participation from users of all experience levels

### 2021-06-14 paper.tex
cleanup repo
a few little updates

--- 2021-06-11 templates/paper.tex
@@
 \newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
-\newcommand{\githubrepository}{\url{https://github.com/GrossfieldLab/article_templates}}
+\newcommand{\githubrepository}{\url{https://github.com/GrossfieldLab/software_community}}
 
@@
 
-In all of these cases, it is necessary to interact with the folks who develop
+In all of these cases, it is necessary to interact with the people who develop
 and support the code. How you go about it has an enormous impact on your
 likelihood of success and how you are viewed, but this is not something that is
-explicitly taught in most places.  {\emph The goal of this work is to reveal the hidden
-curriculum -- what's expected of a software consumer, how to ask for help, how
-to contribute productively to a software community (regardless of whether you
-can write code).}
+explicitly taught in most places.  {\emph The goal of this paper is to reveal
+the hidden curriculum -- what's expected of a software consumer, how to ask for
+help, how to contribute productively to a software community (regardless of
+whether you can write code).}
+
+Asking strangers for help can be intimidating, regardless of your
 
@@
 
-\section{Prerequisites}
+emphasize the idea of a social contract: what do developers owe the users, what
+do users owe developers, what do users owe other users
 
-Training articles should clearly state the target audience and knowledge prerequisites.
-Key prerequisites should be noted in the article abstract to permit readers to rapidly ascertain an articles suitability.
+connect to open science: FAIR principles say developers should want code to be
+reusable. Publishing the code isn't enough, you need to make it usable as well.
+That means support.
 
-\subsection{Background knowledge}
+\section{Target Audience}
 
-Although the authors may imagine a particular career-level (e.g., undergraduate or graduate), given the diversity of disciplinary curricula, it is more important to specify precisely any knowledge prerequisites (e.g., vector calculus, basic thermodynamics).
-
-\subsection{Software/system requirements}
-
-If a particular software or programming environment plays a central role in the article, that should be specified.
-
-\section{Content and links}
-
-A training article may on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
-You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.
+This paper is primarily aimed at junior scientists who are new to performing
+research and inexperienced at asking for help. However, we hope it will be
+valuable to anyone who uses software to do their research and struggles to ask
+for help.
 

### 2021-06-14 paper.tex
some new text

@@
 and support the code. How you go about it has an enormous impact on your
-likelihood of success and how you are viewed, but this is not something that is
-explicitly taught in most places.  {\emph The goal of this paper is to reveal
-the hidden curriculum -- what's expected of a software consumer, how to ask for
-help, how to contribute productively to a software community (regardless of
-whether you can write code).}
+likelihood of success and how you are viewed, and asking complete strangers for
+help is often intimidating, especially the first time you do it. However, the
+best way to do so is rarely taught explicitly --- at best, it's something one
+picks up by example, from fellow lab members or the PI.  {\emph The goal of this
+paper is to reveal the hidden curriculum -- what's expected of a software
+consumer, how to ask for help, how to contribute productively to a software
+community (regardless of whether you can write code).}
 
-Asking strangers for help can be intimidating, regardless of your
-
-There lots of articles on best practices in software development, including open
-source.
-
-emphasize the idea of a social contract: what do developers owe the users, what
-do users owe developers, what do users owe other users
+This article is written from a perspective of academic open-source scientific
+software. Much of what we suggest is universal, but some points --- like the
+reminder that the people providing the support are likely volunteers --- are
+not.  Regardless, we view the interaction between developers and users as an
+informal social contract, where each has obligations and expectations for the
+other. Obviously, every software community is unique, and many have specific
+conventions for interaction, but we hope that the recommendations we make are at
+the very least a good starting point for new users of scientific software.
 
@@
 for help.
+
+\section{Good practices in asking for help}
+
+\section{Obligations of software developers}
+
+This paper has focused on what software consumers should do when asking for
+help.  However, this does not mean that the developers of scientific software
+are without responsibilities. If we want users to behave in a certain way, it
+behooves us to tell them that. Users approaching us in good faith deserve to be
+treated fairly and courteously.
 

### 2021-07-08 paper.tex
more text

@@
 
+\subsection{Try to solve your own problem}
+
+There's absolutely nothing wrong with needing help to figure out how to do
+something with a new piece of software. Perhaps your use case wasn't considered
+in the manual, or you're seeing behavior you don't expect. Perhaps you can't
+even get it to install. Asking for help is very reasonable --- the developers
+want people to use their software, and with that comes an implied obligation to
+support those users. However, keep in mind that the developers of scientific
+software are, for the most part, volunteering their time to provide that
+support, and as such it's important to respect their time.
+
+For this reason, asking the developers for help shouldn't be the \emph{first}
+thing you do. Especially for software with a large community, chances are
+someone else has run into similar difficulties, and you can probably save
+yourself and the developers a lot of time if you look for solutions on your own
+first. Indeed, quickly debugging your use of others' software is one of the
+major skills one develops when learning to do computational research.
+
+The first step is to check the manual. This might seem trivially obvious, but you'd be astonished how many emails developers get that are directly answered in the manual.
+
+This might be an actual standalone
+manual, or part of the web page for the software package, but if it exists it
+should be your first stop. Look for a frequently asked questions list on the web site,
+
+Search the internet for the error message (this works very well for software with thousands of users, less well with smaller packages)
+
 \section{Obligations of software developers}
@@
 \item Cites and acknowledges software appropriately
+\item Treats fellow members and developers with courtesy and respect
 \item Contributes to the community
-\item Treats fellow members and developers with courtesy and respect
 \end{itemize}

### 2021-07-08 paper.tex
a bit more text

@@
 yourself and the developers a lot of time if you look for solutions on your own
-first. Indeed, quickly debugging your use of others' software is one of the
-major skills one develops when learning to do computational research.
+first. Indeed, quickly debugging your use of other people's software is one of
+the major skills one develops when learning to do computational research.
 
-The first step is to check the manual. This might seem trivially obvious, but you'd be astonished how many emails developers get that are directly answered in the manual.
+The first step is to read the error message carefully, if there is one; this
+obviously doesn't apply in all cases, but it's often applicable and far too
+often skipped. It's true that many software packages have cryptic unhelpful
+error messages, but far too often people write for help when the solution is
+right in front of them. If the error message says ``Could not open file
+foo.dat'', it's worth checking whether there is a file called ``foo.dat'' and if
+not figuring out what was supposed to create it.
 
-This might be an actual standalone
-manual, or part of the web page for the software package, but if it exists it
-should be your first stop. Look for a frequently asked questions list on the web site,
+The second step is to check the manual. This might seem trivially obvious, but
+you'd be astonished how many emails developers get that are directly answered in
+the manual. If the manual is short, you can read the whole thing. If it's
+longer, search for relevant-seeming keywords, scan the sections that seem to
+pertain to your error message. If there's an FAQ (frequently asked questions)
+list, check that first.
 
-Search the internet for the error message (this works very well for software with thousands of users, less well with smaller packages)
+Next, search the internet for the error message; this works very well for
+software with thousands of users, where it might very well deserve to be the
+first option. It is less effective with less common packages, but it's an easy
+thing to check, so you'd be foolish to skip it.
+
+Finally, you can consider diving into the code. This decision depends very much
+on the complexity of the package, your skill level, and how urgently you need
+the solution. More often than not, comprehending the whole software package will
+be too time consuming, but there are tricks and shortcuts one can use to at
+least figure out where things might be going wrong. Figuring out where the error
+is occurring can sometimes help you figure out why it happened, so examine the
+stack trace if there is one. Alternatively, you can search the code base for the
+error message, to see what drove the error; it's not uncommon for the comments
+in the code to be clearer and more indicative of what could go wrong than the
+error message itself. To be clear: as an end user, \emph{you are not obligated
+to check the code}, but it can sometimes be a good shortcut to help you either
+solve your problem on your own or give the developers what they need to solve
+it.
 

### 2021-07-13 paper.tex
more text

@@
 and support the code. How you go about it has an enormous impact on your
-likelihood of success and how you are viewed, and asking complete strangers for
+likelihood of success and how you are viewed. Asking complete strangers for
 help is often intimidating, especially the first time you do it. However, the
@@
 
-This article is written from a perspective of academic open-source scientific
-software. Much of what we suggest is universal, but some points --- like the
+This article is written from a the perspective of a developer of academic
+open-source scientific software, based on my personal experience as a developer
+and user. I released my first piece of open source code, wham \cite{WHAM}, a
+tool for analyzing umbrella sampling simulation,  during the summer of 2000.  My
+group released LOOS\cite{Grossfield-2009, LOOS-JCC}, a suite of tools for
+developing new tools to analyze molecular dynamics simulations, in 2008. Over
+the years, I've interacted with hundreds of people who've tried to use one
+package or the other. Most of these interactions have been positive, some
+overwhelmingly so. Others were not. This article is an attempt to distill my
+experiences into a concise set of recommendations.
+
+Much of what I suggest is universal, but some points --- like the
 reminder that the people providing the support are likely volunteers --- are
@@
 the very least a good starting point for new users of scientific software.
-
-connect to open science: FAIR principles say developers should want code to be
-reusable. Publishing the code isn't enough, you need to make it usable as well.
-That means support.
 
@@
 solve your problem on your own or give the developers what they need to solve
-it.
+it. Moreover, if you do solve the problem, it's a good idea to share your
+solution with the developers; most will be grateful to you for your help.
+
+\subsection{Ask for help in the right place}
+
+\subsection{Write a good bug report}
+
+\subsection{Contribute to the community}
+
+\subsection{Treat people with respect}
+
+This should go without saying, but if you're asking for help it behooves you to
+do so respectfully. We're not saying that you need to be obsequious. Rather,
+approach every interaction with respect: respect for people's time, respect for
+their ideas, respect for their identity. If you're asking for help, do so with
+the perspective that the developer is volunteering their time to help you, with
+minimal payoff to them.  If you're reporting a bug, you're almost certainly
+frustrated; try not to take that frustration out on the developers. If you're
+offering a new feature, keep in mind that it may not align with the developers'
+vision for where the software is going.
+
+If you're interacting with fellow users in a support forum, don't mock
+questioners who are less experienced than you --- a question may seem foolish to
+you, but even an uninformed or ill-directed question is an opportunity for
+teaching; one of our goals for this document is that it become a resource to
+educate new users of expected norms without consuming developers' time and
+mental bandwidth.
 
@@
 treated fairly and courteously.
+
+connect to open science: FAIR principles say developers should want code to be
+reusable. Publishing the code isn't enough, you need to make it usable as well.
+That means support.
+
+Treating everyone with courtesy and respect, regardless of their identity, is
+essential if you want the software tool to have a vibrant community surrounding
+it. Ensuring people aren't excluded is the fair and just way to proceed ---
+making sure that racist, sexist, anti-LGBTQ, etc language and behavior is
+unwelcome and unacceptable is simply the moral thing to do (it's also pragmatic
+--- can you really afford to lose talented developers and users because they
+don't feel welcome?) This doesn't just mean use of epithets in messages (though
+that's clearly unacceptable). Keep an eye on how things are named, on jokes
+present in the code or messages, on whose opinions are listened to and who is
+ignored.  Above all, if someone tells you a particular behavior or situation is
+harmful, \emph{believe them}. Something that feels minor (or is just  a joke) to
+you may hit others very differently, and part of respect is understanding and
+accounting for that.
 
@@
 \item Cites and acknowledges software appropriately
+\item Contributes to the community
 \item Treats fellow members and developers with courtesy and respect
-\item Contributes to the community
 \end{itemize}

### 2021-07-14 paper.tex
more text, reference github templates

@@
 
+Once you've determined you need help, the next obvious step is to ask for it.
+How should you go about doing so? Email the developer? Post in a forum? Raise an
+issue on GitHub? Asking the question in the right venue will greatly increase
+your chances of getting a timely answer. Step 1 is to check the documentation:
+look at the manual, check the README on the GitHub repository, check the
+software website, etc --- most of the time, at least one of these sources will
+say how they prefer to receive bug reports, feature requests, etc. Actually,
+there's a step 0: make sure you're looking at the right piece of software. This
+might seem obvious, but I've received multiple requests for help with software I
+had nothing to do with; once it took 8 emails back and forth before I figured
+this out (but more on that problem later).
+
+Asking in the right place is particularly important if the software is complex
+or has a large user community --- there may be separate forums for bug reports,
+usage help, and feature request, and asking in the wrong place will make it
+unlikely the right people will see your question. Moreover, the first impression
+you'll make is that you're someone who can't be bothered to read the
+instructions, which will make them less inclined to help you.
+
 \subsection{Write a good bug report}
+
+GitHub issue templates \cite{github-templates}
 
@@
 you, but even an uninformed or ill-directed question is an opportunity for
-teaching; one of our goals for this document is that it become a resource to
-educate new users of expected norms without consuming developers' time and
-mental bandwidth.
+teaching. That said, we understand that answering the same questions repeatedly
+is frustrating and exhausting; one of our goals for this document is that it
+become a resource to educate new users of expected norms without consuming
+developers' time and mental bandwidth.
 

### 2021-07-27 paper.tex
a little more text and a ref

@@
 
+The next step in the process is to actually communicate your issue with the
+developers, generally in the form of a bug report.  Doing so effectively
+requires striving for two somewhat contradictory goals: you must give the
+developer all of the relevant information about the problem, as concisely as
+possible.
+
+The reason these two tasks are complicated is that, as a user, you often don't know what information is relevant.
+
+rubber duck debugging \cite{Thomas-1999}-- the process of writing a good bug
+report is analogous to solving the problem. Often, you can't really do it,
+because if you knew what was necessary, you'd be able to solve your problem.
+
 GitHub issue templates \cite{github-templates}

### 2021-08-11 paper.tex
more text

--- 2021-07-14 paper.tex
@@
 \subsection{Try to solve your own problem}
+\label{ss:yourself}
 
@@
 
-GitHub issue templates \cite{github-templates}
+The next step in the process is to actually communicate your issue with the
+developers, generally in the form of a bug report.  Doing so effectively
+requires striving for two somewhat contradictory goals: you must give the
+developer all of the relevant information about the problem, as concisely as
+possible.
+
+Taking the first part first: this is hard, because as a user you're usually not
+an expert, \emph{so you don't know what's relevant}. The only way to know for
+sure what is and isn't relevant is to develop a mental model for the problem,
+which requires effort. Still, as pointed out above, your prior efforts to solve
+the problem yourself (see \ref{ss:yourself}) will help you here.  Moreover, you
+may find that the act of writing the note asking for help may in itself let you
+solve your problem. The process is analogous to rubber duck debugging
+\cite{Thomas-1999}, in that the act of figuring out how to describe a problem
+clearly can lead you to a solution.
+
+Moreover, many developers will help you with the process. For example,
+GitHub-based projects have the option to set up issue templates
+\cite{github-templates}; as the name implies, these templates contain a series
+of questions one should answer when submitting a new issue. Although they are
+generally customized for each project, these questions are also a good baseline
+for the kinds of information one should supply when asking for support
+elsewhere. For example, for a bug report one should
+
+\begin{itemize}
+
+    \item Describe the bug, including the expected behavior and how the current
+    behavior differs.
+
+    \item Give steps necessary to reproduce the bug.
+
+    \item Describe the platform you're running on (e.g. the operating system for
+    a conventional computer, hardware if it's a tablet, etc.). If the problem is
+    related to building or installing the software, describe the build system
+    (e.g. compiler version) or package manager (e.g. Conda\cite{ANACONDA}).
+
+\end{itemize}
+
+For a feature request, you would also describe what you hope to accomplish and
+how you'd like to see it work, as well as any current workaround you've
+developed.
 
@@
 
-Alan Grossfield serves as a consultant to two companies, Moderna Therapeutics
+Alan Grossfield serves as a consultant to two companies, Moderna, Inc.
 and Atelerix Life Sciences.

### 2021-08-16 paper.tex
more text
lots more text

@@
 pertain to your error message. If there's an FAQ (frequently asked questions)
-list, check that first.
+list, check that first; asking an FAQ is likely to lead to grumbling.
 
@@
 it. Moreover, if you do solve the problem, it's a good idea to share your
-solution with the developers; most will be grateful to you for your help.
+solution with the developers; most will be grateful to you for your help, and
+they have comments or improvements suggested improvements. If the package is
+hosted on GitHub, consider opening a pull request so your solution can be merged
+into the main branch.
 
@@
 
+    \item Give the version of the code you're using.
+
     \item Describe the platform you're running on (e.g. the operating system for
@@
 
-For a feature request, you would also describe what you hope to accomplish and
-how you'd like to see it work, as well as any current workaround you've
-developed.
+Sometimes, the developer will be able to answer your question based on the first
+email. However, their reply will often contain questions or suggestions for
+things to test. \emph{It is crucial that you read this message carefully, and
+answer all of the questions to the best of your ability when you reply.} As a
+rule, these questions aren't asked idly; the person providing support is trying
+to help you! Far too often, users don't read the reply email carefully, don't
+answer the questions, try only the first suggestion, or refuse to indicate what
+it is they were actually doing. If you do this, the best case is that it will
+take longer for you to get to a solution. More likely, you will frustrate the
+person trying to help you, and they may become reluctant to work on your issue.
+
+Feature requests are similar: you need to describe what you hope to accomplish
+and how you'd like to see it work, as well as any current workaround you've
+developed. Keep in mind that many features sound easy, but there can be deep
+design issues that make their implementation difficult. Moreover, it is possible
+that the developer won't be enthusiastic about your idea, even if it is doable
+--- for example, it might not coincide with their vision for the software. Or,
+as discussed above, academic developers are generally not compensated for their
+work (obtaining NIH funds to support software development is extremely
+difficult) and simply might not have the time to work on new features.
 
@@
 approach every interaction with respect: respect for people's time, respect for
-their ideas, respect for their identity. If you're asking for help, do so with
-the perspective that the developer is volunteering their time to help you, with
-minimal payoff to them.  If you're reporting a bug, you're almost certainly
-frustrated; try not to take that frustration out on the developers. If you're
-offering a new feature, keep in mind that it may not align with the developers'
-vision for where the software is going.
+their ideas, respect for their identity (including among other things race,
+ethnicity, country, sex or gender, or career stage). If you're asking for help,
+do so with the perspective that the developer is volunteering their time to help
+you, with minimal payoff to them.  If you're reporting a bug, you're almost
+certainly frustrated; try not to take that frustration out on the developers. If
+you're offering a new feature, keep in mind that it may not align with the
+developers' vision for where the software is going.
 
@@
 This paper has focused on what software consumers should do when asking for
-help.  However, this does not mean that the developers of scientific software
-are without responsibilities. If we want users to behave in a certain way, it
-behooves us to tell them that. Users approaching us in good faith deserve to be
-treated fairly and courteously.
+help.  I suppose it could even be read as a list of complaints about bad
+behavior by users (though that's not how I intend it). However, each behavior
+described above has a corresponding expectation for the developers.
+
+However, this does not mean that the developers of scientific
+software are without responsibilities. If we want users to behave in a certain
+way, it behooves us to tell them that. Users approaching us in good faith
+deserve to be treated fairly and courteously.
 
@@
 accounting for that.
+
+Moreover, it's not enough for you as a developer to treat people with respect.
+To the extent that there's a community around your software, you must attempt to
+make sure that community is respectful; as developer, your words and deeds carry
+significant weight. If you call out bad behavior (or even marginal behavior),
+that will go a long way towards setting a good tone for your community. If you
+let it slide, chances are others will assume you implicitly endorse it.
 
@@
 % We suggest you preserve this comment:
-For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.
+For a more detailed description of contributions from the community and others, see the GitHub issue tracking and Changelog at \githubrepository.
 
@@
 %%%%%%%
-This work supported in part by NIH R21GM138970 to AG.
+This work supported in part by NIH R21GM138970 to Alan Grossfield.
 

### 2021-08-17 paper.tex
more text

@@
 
-    \item Give the version of the code you're using.
+    \item Give the version of the code you're using. If you've changed anything,
+    say what you've changed. If there are compile-time options, say what you
+    chose.
 
@@
 person trying to help you, and they may become reluctant to work on your issue.
+In my experience, few things are more frustrating than trying to help someone
+who won't help me help them.
 
@@
 difficult) and simply might not have the time to work on new features.
-
-\subsection{Contribute to the community}
 
@@
 developers' time and mental bandwidth.
+
+\subsection{Contribute to the community}
+
+Once you've become reasonably competent with a piece of software, it's worth
+looking for ways to contribute back. This can take many forms, not all of which
+involve writing code (though of course contributing bug fixes or new features is
+extremely valuable). You can work on the documentation --- if you struggled to
+understand something but eventually figured it out, consider contributing a new
+paragraph of text; many packages hosted on GitHub have the documentation in the
+same repository, so the same pull request mechanism can be used, but even if the
+docs are separate you could write a new paragraph or two and send them to the
+developer. If you developed a toy example as you were figuring out how to use
+the software, consider sharing it. If one of the provided examples had to be
+tweaked to work, share the tweaks. If the software has an online forum (e.g.
+mailing list, GitHub discussions or issues, etc), you might be able to
+contribute by suggestion solutions to problems from less experienced users. However, unlike the other recommentations, this one isn't mandatory --- it's a ``nice to have'' rather than a ``must have''.
 
@@
 \item Encourages participation from users of all experience levels
-\item Encourages respectful treatment of all community members, and calls out disrespectful behavior
+\item Encourages respectful treatment of all community members, and calls out bad behavior
 \end{itemize}
@@
 
-Alan Grossfield serves as a consultant to two companies, Moderna, Inc.
+Alan Grossfield serves as a consultant to two companies: Moderna, Inc.
 and Atelerix Life Sciences.

### 2021-08-19 paper.tex
more text -- at least something for every section, I think

@@
 This paper is primarily aimed at junior scientists who are new to performing
-research and inexperienced at asking for help. However, we hope it will be
-valuable to anyone who uses software to do their research and struggles to ask
-for help.
+computational research and inexperienced at asking for help. However, we hope it
+will be valuable to anyone who uses software to do their research and struggles
+to ask for help.
 
@@
 mailing list, GitHub discussions or issues, etc), you might be able to
-contribute by suggestion solutions to problems from less experienced users. However, unlike the other recommentations, this one isn't mandatory --- it's a ``nice to have'' rather than a ``must have''.
+contribute by suggestion solutions to problems from less experienced users.
+However, unlike the other recommentations, this one isn't mandatory --- it's a
+``nice to have'' rather than a ``must have''.
 
@@
 behavior by users (though that's not how I intend it). However, each behavior
-described above has a corresponding expectation for the developers.
+described above has a corresponding expectation for the developers. If we expect
+the users to try to solve their own problem, it behooves us to write good
+document, give examples, and (most importantly) keep them up to date. If we want
+to be contacted via GitHub only, we need to make that information clear and easy
+to find. If we are frustrated by bug reports that don't give us the information
+we need, we should tell people what kinds of information we need (e.g. by
+providing bug report templates). In order to receive proper credit, we need to
+document the appropriate way to cite the software. All of this information needs
+to be clearly written out and easy to find.
 
-However, this does not mean that the developers of scientific
-software are without responsibilities. If we want users to behave in a certain
-way, it behooves us to tell them that. Users approaching us in good faith
-deserve to be treated fairly and courteously.
+Doing these things is valuable from a practical perspective; making it easy for
+people to interact with us efficiently, even if it costs us some effort up
+front, will make those interactions more pleasant. Even setting that aside, it's
+the right way to do science --- FAIR principles for software development
+\cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
+That means documenting it well and supporting people who try to use it.
 
-connect to open science: FAIR principles say developers should want code to be
-reusable. Publishing the code isn't enough, you need to make it usable as well.
-That means support.
+Users approaching us in good faith deserve to be treated fairly and courteously.
+Treating everyone with courtesy and respect, regardless of their identity, is
+essential if you want to build a vibrant effective community.  If the community
+is just you, the sole developer, then the solution is obvious --- when you
+interact with users, do so in ways that make them feel valued and respected.
 
-Treating everyone with courtesy and respect, regardless of their identity, is
-essential if you want the software tool to have a vibrant community surrounding
-it. Ensuring people aren't excluded is the fair and just way to proceed ---
-making sure that racist, sexist, anti-LGBTQ, etc language and behavior is
-unwelcome and unacceptable is simply the moral thing to do (it's also pragmatic
---- can you really afford to lose talented developers and users because they
-don't feel welcome?) This doesn't just mean use of epithets in messages (though
-that's clearly unacceptable). Keep an eye on how things are named, on jokes
-present in the code or messages, on whose opinions are listened to and who is
-ignored.  Above all, if someone tells you a particular behavior or situation is
-harmful, \emph{believe them}. Something that feels minor (or is just a joke) to
-you may hit others very differently, and part of respect is understanding and
-accounting for that.
+As the user and developer community grows, other people will get involved in
+supporting users. It's still your obligation as a developer to maintain those
+standards.  Making sure that racist, sexist, anti-LGBTQ, etc language and
+behavior is unwelcome and unacceptable is simply the moral thing to do (it's
+also pragmatic --- can you really afford to lose talented developers and users
+because they don't feel welcome?) This doesn't just mean use of epithets in
+messages (though that's clearly unacceptable). Keep an eye on how things are
+named, on jokes present in the code or messages, on whose opinions are listened
+to and who is ignored.  Above all, if someone tells you a particular behavior or
+situation is harmful, \emph{believe them}.  Something that feels minor (or is
+just a joke) to you may hit others very differently, and part of respect is
+understanding and accounting for that.
 
-Moreover, it's not enough for you as a developer to treat people with respect.
-To the extent that there's a community around your software, you must attempt to
-make sure that community is respectful; as developer, your words and deeds carry
-significant weight. If you call out bad behavior (or even marginal behavior),
-that will go a long way towards setting a good tone for your community. If you
-let it slide, chances are others will assume you implicitly endorse it.
+Along the same lines, a developer who berates a user for asking a stupid
+question, or is rude and dismissive in a technical discussion about a new
+feature, is actively damaging the community, \emph{regardless of the technical
+quality of their contributions.} It's not enough for you as a developer to treat
+people with respect.  To the extent that there's a community around your
+software, you must attempt to make sure that community is respectful; as
+developer, your words and deeds carry significant weight. If you call out bad
+behavior (or even marginal behavior), that will go a long way towards setting a
+good tone for your community. If you let it slide, chances are others will
+assume you implicitly endorse it.
 

### 2021-08-28 paper.tex
incorporated some of Tod's ideas

@@
 are not programmers, and must rely on programs written and maintained by others.
-As a result, a crucial but often neglected part of a scientist's training is
+A crucial but often neglected part of a scientist's training is
 learning how to use new tools, and how to exist as part of a community of users.
@@
 often means learning to use new programs or new aspects of large packages. This,
-combined with the complex nature of scientific software (and the often low
-quality of the associated documentation) means we have to ask for help.
-Sometimes it's because we don't know how to accomplish a specific task. Others,
-it's because there's a feature we need that isn't implemented. Inevitably, there
-are bugs.
+combined with the complex nature of scientific software (and often the missing
+or inaccurate documentation) means we have to ask for help.  Sometimes it's
+because we don't know how to accomplish a specific task. Others, it's because
+there's a feature we need that isn't implemented. Inevitably, there are bugs.
 
@@
 and support the code. How you go about it has an enormous impact on your
-likelihood of success and how you are viewed. Asking complete strangers for
-help is often intimidating, especially the first time you do it. However, the
-best way to do so is rarely taught explicitly --- at best, it's something one
-picks up by example, from fellow lab members or the PI.  {\emph The goal of this
+likelihood of success and how you are viewed. Asking complete strangers for help
+is often intimidating, especially the first time you do it. However, the best
+way to do so is rarely taught explicitly --- at best, it's something one picks
+up by example, from fellow lab members or your PI.  {\emph The goal of this
 paper is to reveal the hidden curriculum -- what's expected of a software
@@
 pertain to your error message. If there's an FAQ (frequently asked questions)
-list, check that first; asking an FAQ is likely to lead to grumbling.
+list, check that first; asking an FAQ, especially in a public forum, is likely
+to lead to grumbling.
 
@@
 thing to check, so you'd be foolish to skip it.
+
+One valuable strategy is to construct a minimal set of circumstances that causes
+the problem. Even if you initially encountered the problem while analyzing a 2
+TB data set, that's not going to be very helpful for diagnosing the problem (and
+you won't easily be able to share the data set with the developers so they can
+diagnose it). Can you reproduce the problem using only a small portion of the
+data? Is the problem present with some kinds of data (e.g. some frames of the
+trajectory) but not others, and if so what is different about those frames? Can
+you make the problem appear and disappear with small changes in the settings?
+Testing these things out may lead you to a solution, but even if they don't
+they'll help you describe the problem more precisely to the developers.
 
@@
     related to building or installing the software, describe the build system
-    (e.g. compiler version) or package manager (e.g. Conda\cite{ANACONDA}).
+    (e.g. compiler version) or package manager (e.g. Conda\cite{CONDA}, Homebrew \cite{HOMEBREW}).
 
@@
 understand something but eventually figured it out, consider contributing a new
-paragraph of text; many packages hosted on GitHub have the documentation in the
+paragraph of text. Many packages hosted on GitHub have the documentation in the
 same repository, so the same pull request mechanism can be used, but even if the
@@
 mailing list, GitHub discussions or issues, etc), you might be able to
-contribute by suggestion solutions to problems from less experienced users.
-However, unlike the other recommentations, this one isn't mandatory --- it's a
-``nice to have'' rather than a ``must have''.
+contribute by answering questions from less experienced users.
+
+Unlike the other recommentations, this one isn't mandatory --- it's a ``nice to
+have'' rather than a ``must have''. However, there are a number of advantages to
+doing it. First and foremost, you'll be making a package that's valuable to you
+stronger, which will help you with your own scientific projects. However,
+participating productively in a software community is also a good way to build
+your scientific communication skills. Finally, helping others is an excellent
+way for a relatively junior scientist to build their reputation in the
+community, which in turn can lead to future collaborations or even job
+opportunities. At the very least, it's always valuable when a lot of people in
+your field have positive associations with your name.
 
@@
 assume you implicitly endorse it.
+
+If the justice perspective doesn't persuade you, consider the practical
+consequences. Bad behavior on the part of developers of your project (or your
+community in general) will drive users and developers away from you. If
+interacting with your community is unpleasant, folks will either use alternative
+codes or use your code but not cite you. The next person with an idea for an
+innovative new feature will add it to a different package, or won't do it at
+all (hurting the whole community).
+Ask yourself: can your project really afford to drive away talented hard-working people eager to help?
+Is it worth is, just so a few people don't have to engage with basic courtesy?
 

### 2021-08-29 paper.tex
acknowledged Tod
incorporated Justin Lemkul's suggestions

@@
 way to do so is rarely taught explicitly --- at best, it's something one picks
-up by example, from fellow lab members or your PI.  {\emph The goal of this
+up by example, from fellow lab members or your PI.  \emph{The goal of this
 paper is to reveal the hidden curriculum -- what's expected of a software
@@
 
+Along the same lines, it is often valuable to construct a fake or minimal data
+set where the correct answer is known (or easily computed outside the program).
+This is particularly important when your concern is for the program's
+correctness. This is particularly important when your concern is that the
+program is producing incorrect answers due to a coding error or assumptions
+incompatible with the input data.
+
 Finally, you can consider diving into the code. This decision depends very much
@@
 
-    \item Give steps necessary to reproduce the bug.
+    \item Give steps necessary to reproduce the bug. Include any input data
+    required.
 
@@
 their ideas, respect for their identity (including among other things race,
-ethnicity, country, sex or gender, or career stage). If you're asking for help,
-do so with the perspective that the developer is volunteering their time to help
-you, with minimal payoff to them.  If you're reporting a bug, you're almost
-certainly frustrated; try not to take that frustration out on the developers. If
-you're offering a new feature, keep in mind that it may not align with the
-developers' vision for where the software is going.
+ethnicity, country, native language, sex or gender, or career stage). If you're
+asking for help, do so with the perspective that the developer is volunteering
+their time to help you, with minimal payoff to them.  If you're reporting a bug,
+you're almost certainly frustrated; try not to take that frustration out on the
+developers. If you're offering a new feature, keep in mind that it may not align
+with the developers' vision for where the software is going.
 
@@
 understanding and accounting for that.
+
+It's also important to recognize that even while English has become the default
+language for science, not everyone is a native speaker, which means that some
+jargon, abbreviations, and idioms may be incomprehensible to a fair portion of
+your community. Moreover, an awkwardly worded question often reflects lack of
+familiarity with the language rather than laziness or ignorance. It's important
+to keep accessibility in mind while creating documentation or interacting with
+users in a forum.
 
@@
 %%%%%%%%%%%%%%%
+Dr. Tod D. Romo reviewed the document for style and content. Dr. Justin Lemkul
+made numerous helpful suggestions, particularly regarding language
+accessibility.
 

### 2021-08-30 paper.tex
caught a typo
wording tweak

@@
 computational research and inexperienced at asking for help. However, we hope it
-will be valuable to anyone who uses software to do their research and struggles
-to ask for help.
+will be valuable to anyone who uses software written by others to do their
+research.
 
@@
 Ask yourself: can your project really afford to drive away talented hard-working people eager to help?
-Is it worth is, just so a few people don't have to engage with basic courtesy?
+Is it worth it, just so a few people don't have to engage with basic courtesy?
 

### 2021-08-30 paper.tex
Dave's suggestions

@@
 set where the correct answer is known (or easily computed outside the program).
-This is particularly important when your concern is for the program's
-correctness. This is particularly important when your concern is that the
-program is producing incorrect answers due to a coding error or assumptions
-incompatible with the input data.
+This is particularly important when your concern is that the program is
+producing incorrect answers due to a coding error or assumptions incompatible
+with the input data. The more you precisely you can narrow down what's going
+wrong, the easier the problem will be to diagnose.
 
@@
 standards.  Making sure that racist, sexist, anti-LGBTQ, etc language and
-behavior is unwelcome and unacceptable is simply the moral thing to do (it's
-also pragmatic --- can you really afford to lose talented developers and users
-because they don't feel welcome?) This doesn't just mean use of epithets in
-messages (though that's clearly unacceptable). Keep an eye on how things are
-named, on jokes present in the code or messages, on whose opinions are listened
-to and who is ignored.  Above all, if someone tells you a particular behavior or
-situation is harmful, \emph{believe them}.  Something that feels minor (or is
-just a joke) to you may hit others very differently, and part of respect is
-understanding and accounting for that.
+behavior is unwelcome and unacceptable is simply the moral thing to do.  This
+doesn't just mean use of epithets in messages (though that's clearly
+unacceptable). Keep an eye on how things are named, on jokes present in the code
+or messages, on whose opinions are listened to and who is ignored.  Above all,
+if someone tells you a particular behavior or situation is harmful,
+\emph{believe them}.  Something that feels minor (or is just a joke) to you may
+hit others very differently, and part of respect is understanding and accounting
+for that.
 
@@
 codes or use your code but not cite you. The next person with an idea for an
-innovative new feature will add it to a different package, or won't do it at
-all (hurting the whole community).
-Ask yourself: can your project really afford to drive away talented hard-working people eager to help?
-Is it worth it, just so a few people don't have to engage with basic courtesy?
+innovative new feature will add it to a different package, or won't do it at all
+(hurting the whole community). Ask yourself: can your project really afford to
+drive away talented hard-working people eager to help? Is it worth it, just so a
+few people don't have to engage with basic courtesy?
 
@@
 %%%%%%%%%%%%%%%
-Dr. Tod D. Romo reviewed the document for style and content. Dr. Justin Lemkul
-made numerous helpful suggestions, particularly regarding language
-accessibility.
+Dr. Tod D. Romo reviewed the document for style and content. Dr. Justin A.
+Lemkul made numerous helpful suggestions, particularly regarding language
+accessibility. Dr. David H. Mathews provided several editorial suggestions.
 

### 2021-08-31 paper.tex
more text

@@
 obviously doesn't apply in all cases, but it's often applicable and far too
-often skipped. It's true that many software packages have cryptic unhelpful
+often skipped. It's true that many software packages have cryptic or unhelpful
 error messages, but far too often people write for help when the solution is
@@
 
-The second step is to check the manual. This might seem trivially obvious, but
-you'd be astonished how many emails developers get that are directly answered in
-the manual. If the manual is short, you can read the whole thing. If it's
-longer, search for relevant-seeming keywords, scan the sections that seem to
-pertain to your error message. If there's an FAQ (frequently asked questions)
+The second step is to check the manual. This too might seem trivially obvious,
+but you'd be astonished how many emails developers get that are directly
+answered in the manual. If the manual is short, you can read the whole thing. If
+it's longer, search for relevant-seeming keywords, scan the sections that seem
+to pertain to your error message. If there's an FAQ (frequently asked questions)
 list, check that first; asking an FAQ, especially in a public forum, is likely
@@
 first option. It is less effective with less common packages, but it's an easy
-thing to check, so you'd be foolish to skip it.
+thing to check, so you'd be foolish to skip it.  Search for the whole text of
+the error (excluding any verbiage that looks like it depends on the compiler,
+etc), with and without the package name. The latter can be important in the case
+where the problem isn't with the package per se, but with something else in your
+setup.
 
@@
 Lemkul made numerous helpful suggestions, particularly regarding language
-accessibility. Dr. David H. Mathews provided several editorial suggestions.
+accessibility. Dr. David H. Mathews provided several editorial suggestions. Dr.
+Jeffrey Lewis had suggestions regarding searching for help.
 

### 2021-08-31 paper.tex
fixed typo

@@
 the users to try to solve their own problem, it behooves us to write good
-document, give examples, and (most importantly) keep them up to date. If we want
-to be contacted via GitHub only, we need to make that information clear and easy
-to find. If we are frustrated by bug reports that don't give us the information
-we need, we should tell people what kinds of information we need (e.g. by
-providing bug report templates). In order to receive proper credit, we need to
-document the appropriate way to cite the software. All of this information needs
-to be clearly written out and easy to find.
+documentation, give examples, and (most importantly) keep them up to date. If we
+want to be contacted via GitHub only, we need to make that information clear and
+easy to find. If we are frustrated by bug reports that don't give us the
+information we need, we should tell people what kinds of information we need
+(e.g. by providing bug report templates). In order to receive proper credit, we
+need to document the appropriate way to cite the software. All of this
+information needs to be clearly written out and easy to find.
 

### 2022-01-05 response-letter.tex
major changes suggested by reviewers and wrote response letter

\documentclass[11pt]{article}

\usepackage[numbers,super,sort&compress,comma]{natbib}
\usepackage[pdftex,colorlinks=true,urlcolor=black,filecolor=black,linkcolor=black,citecolor=black,breaklinks=true]{hyperref}
\usepackage{hypernat}
\usepackage[pdftex]{graphicx}
\usepackage{palatino} % make fonts Helvetica for sans, palatino for serif
\usepackage[usenames,dvipsnames]{color}

\newcommand{\response}[1]{\textcolor{Mahogany}{\textbf{#1}}}

\begin{document}

I thank the reviewers for their kind and constructive comments. I hope they agree that incorporating their suggestions has significantly improved the manuscript. I've listed the comments requiring responses below, with my response colored \response{in mahogony}.

\section{Reviewer A}

1) My biggest concern is simply that the people who ought to read it,
won't do so (even if directed) because it is longer than what users
typical want to read when they are in search of a fix to their
problem. In this the article shares the same shortcoming with the
well-known online essays [1] and [2].

I'd like a condensed summary of about one paragraph (maybe as a
box) that I can just paste in an email or put on a website,
together with a link to the full version.

\response{I have added references to the Raymond and Tatham essays -- I'm
somewhat embarrassed I didn't think to add them myself, since they clearly
influenced my thinking when I read them. I would argue that the checklists make
a good short punchy summary, as the reviewer suggested. However, I agree having
a plain-text version wouldn't hurt, so I've added a ``Summary'' paragraph before
the checklists.}

2) Does the author make a distinction between software that is
advertised as useable (e.g. through project websites, presence of
forums, issue tracker, ...) and software that has been made
available (archive, repository, supplementary material)?

Can and should a user assume that a developer of the latter will
provide any support?

Perhaps this could be made clearer.

My personal view is that FAIR principles of research make it
essential that code is published together with the research (code
dump with a license). However, for code dumps I wouldn't require
authors to respond to more than the most fundamental inquiries
(e.g., to respond when bugs are found that question the integrity
of the published research). If more is demanded of authors then
that disincentivizes publishing the code with the research.

On the other hand, if code is advertised (presumably with the
expectation of citations or some other benefit to the developer)
then I agree with the author that developers owe their users some
of their time.

\response{My focus was mostly on code that is intended to be used. I've added a
paragraph at the end of the section for developers laying out my thoughts on
this distinction.}

3) "Please open a PR" (or merge request) is a phrase that can strike
fear in the heart of a novice. It might be helpful to demystify
this important part of collaborative software development. Maybe
the a PLoS Comp Biol article [5] or something similar that touches
upon basic software engineering that could provide more guidance.

\response{I agree, and I'd argue that the relevance here is that it falls to the
developers to provide a less intimidating way to contribute. I'd rather not talk
about specific software engineering technologies here, though, just because I
feel like it's a side topic I couldn't do justice to.}

4) It might be useful to prepare the new community member that
conversations — although respectfule — are often to the point and
that most developers won't sugar-coat their words: if something is
wrong, it will be said just like that because ultimately developers
are concerned with maintaining the health of the code base and the
project and have to be clear about what is acceptable.

New users should not feel offended by the phrase "Feel free to work
on it, submit a PR, and we can discuss there." or similar. Unless
you pay the developers to implement what you want, they will likely
encourage you to do most of the work yourself and then spend some
of their scarce time guiding you to a solution.

\response{This is a good point (although I'd argue that developers ought to be a bit more courteous than that). I added a couple of sentences at the end of the bug report section, appended to the paragraph about feature requests.}

5) Pointing out how much one can learn by interacting with experienced
developers is important. This is a chance to learn from the best
people in the field.

Furthermore, the networks that you build as part of a software
project might last you a life time.

\response{I added a sentence mentioning this to the section on contributing to the community.}

6) I appreciate that the article primarily originates in the
experience of the author. I would nevertheless find it useful if,
where possible, other sources could also be cited as validation and
to show that indeed these are broadly shared values.

I found the following references useful but I am emphatically not
implying that they must be included in the article.

Ref [3] has a good section on the importance of communities.

\response{I'm embarrassed to admit I hadn't read either of the suggested papers.
I've now read them and added references to both.}

\section{Reviewer B}

1) This point is already implicitly transparent throughout the paper, but it
might be useful to state this more explicitly right in Section 3.1 as it
(partially) motivates everything that follows: The easier you make it for the
developer to narrow down/reproduce the issue the higher the chance they will
help you. I think it's worth pointing out that developers are usually handling
multiple issues at the same time and typically tend to assign lower priority to
issues that are harder and time-consuming.

\response{This is a very good point. I added a new paragraph under the description of a bug report.}

2) In section 3.5 it is stated: "Once you’ve become reasonably competent with a
piece of software, it’s worth looking for ways to contribute back. [...] You can
work on the documentation". This is a wonderful suggestion. However, I'm afraid
that stated this way, users that don't feel competent enough will be
hesitant/justified in contributing. I believe proposing small changes to the
documentation often does not require competency. On the contrary, devs and
experienced users typically take for granted a lot of things that might be
missing from the docs. My point is that it might be a good idea to rephrase a
bit to encourage even first users to propose changes to the docs as the best way
to give back some of the time the devs/community has dedicated to them.

\response{This is a very good point. The new version begins ``Once you've begun using a piece of software, it's worth looking for ways to
contribute back.  This can take many forms, not all of which involve writing
code (though of course contributing bug fixes or new features is extremely
valuable). You don't have to be particularly expert -- sometimes the best
documentation comes from the inexperienced or newly experienced users, who
vividly remember not understanding something in the manual and can suggest
revisions that would make it clearer. ''}

\section{Reviewer C}

\response{Thank you for your kind words!}

\end{document}

### 2022-01-05 paper.tex
updated COI
major changes suggested by reviewers and wrote response letter

@@
 The next step in the process is to actually communicate your issue with the
-developers, generally in the form of a bug report.  Doing so effectively
-requires striving for two somewhat contradictory goals: you must give the
-developer all of the relevant information about the problem, as concisely as
-possible.
+developers, generally in the form of a bug report.  The challenges in writing an
+effective bug report have been discussed before in the context of free software
+\cite{Raymond, Tatham}.  Doing so requires striving for two somewhat
+contradictory goals: you must give the developer all of the relevant information
+about the problem, as concisely as possible.
 
@@
 
+Following this format (or something similar) makes it much easier for the
+developer to diagnose your problem, and the easier you make it the more likely
+they are to solve your problem quickly. While it's not ideal behavior, busy
+developers (particularly those for whom software support is a side activity) may
+shy away from starting work on a problem that looks hard or uninviting, while a
+clear concise description of the problem is more likely to spur immediate
+action.
+
 Sometimes, the developer will be able to answer your question based on the first
@@
 developed. Keep in mind that many features sound easy, but there can be deep
-design issues that make their implementation difficult. Moreover, it is possible
-that the developer won't be enthusiastic about your idea, even if it is doable
---- for example, it might not coincide with their vision for the software. Or,
-as discussed above, academic developers are generally not compensated for their
-work (obtaining NIH funds to support software development is extremely
-difficult) and simply might not have the time to work on new features.
+design issues or technical nuances that make their implementation difficult.
+Moreover, it is possible that the developer won't be enthusiastic about your
+idea, even if it is doable --- for example, it might not coincide with their
+vision for the software. Or, as discussed above, academic developers are
+generally not compensated for their work (obtaining NIH funds to support
+software development is extremely difficult) and simply might not have the time
+to work on new features. For these reasons, do not be surprised or insulted if
+the developers suggest you work on a particular solution yourself and submit a
+pull request. This doesn't mean they dislike your idea, just that they're busy -- they'll help guide you as best they can, but they can't do the work themselves.
 
@@
 
-Once you've become reasonably competent with a piece of software, it's worth
-looking for ways to contribute back. This can take many forms, not all of which
-involve writing code (though of course contributing bug fixes or new features is
-extremely valuable). You can work on the documentation --- if you struggled to
-understand something but eventually figured it out, consider contributing a new
-paragraph of text. Many packages hosted on GitHub have the documentation in the
-same repository, so the same pull request mechanism can be used, but even if the
-docs are separate you could write a new paragraph or two and send them to the
-developer. If you developed a toy example as you were figuring out how to use
-the software, consider sharing it. If one of the provided examples had to be
+Once you've begun using a piece of software, it's worth looking for ways to
+contribute back.  This can take many forms, not all of which involve writing
+code (though of course contributing bug fixes or new features is extremely
+valuable). You don't have to be particularly expert -- sometimes the best
+documentation comes from the inexperienced or newly experienced users, who
+vividly remember not understanding something in the manual and can suggest
+revisions that would make it clearer.  Many packages hosted on GitHub have the
+documentation in the same repository, so the same pull request mechanism can be
+used, but even if the docs are separate (or if you find the idea of a pull
+request intimidating) you could write a new paragraph or two and send them to
+the developer. If you developed a toy example as you were figuring out how to
+use the software, consider sharing it. If one of the provided examples had to be
 tweaked to work, share the tweaks. If the software has an online forum (e.g.
@@
 
-Unlike the other recommentations, this one isn't mandatory --- it's a ``nice to
+Unlike the other recommendations, this one isn't mandatory --- it's a ``nice to
 have'' rather than a ``must have''. However, there are a number of advantages to
 doing it. First and foremost, you'll be making a package that's valuable to you
-stronger, which will help you with your own scientific projects. However,
-participating productively in a software community is also a good way to build
-your scientific communication skills. Finally, helping others is an excellent
-way for a relatively junior scientist to build their reputation in the
-community, which in turn can lead to future collaborations or even job
-opportunities. At the very least, it's always valuable when a lot of people in
-your field have positive associations with your name.
+stronger, which will help you with your own scientific projects, while getting
+feedback from leading experts in the field. However, participating productively
+in a software community is also a good way to build your scientific
+communication skills. Finally, helping others is an excellent way for a
+relatively junior scientist to build their reputation in the community, which in
+turn can lead to future collaborations or even job opportunities. At the very
+least, it's always valuable when a lot of people in your field have positive
+associations with your name.
 
@@
 
-Doing these things is valuable from a practical perspective; making it easy for
-people to interact with us efficiently, even if it costs us some effort up
-front, will make those interactions more pleasant. Even setting that aside, it's
-the right way to do science --- FAIR principles for software development
+Doing these things is valuable from a practical perspective, as discussed
+elsewhere \cite{Heister-2013, Laayouni-2016}; making it easy for people to
+interact with us efficiently, even if it costs us some effort up front, will
+make those interactions more pleasant. Even setting that aside, it's the right
+way to do science --- FAIR principles for software development
 \cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
@@
 few people don't have to engage with basic courtesy?
+
+Sometimes, code is published that is not particularly intended for use. For
+example, it is good practice to release the scripts used to analyze data in
+connection with the paper publishing the data, but many of these scripts are
+one-offs, and the authors have little incentive to polish them to the point of
+general useability. Rather, the scripts are essentially there as documentation,
+not turn-key programs for others to use. In this case, the authors have less of
+an obligation to provide support, although at a minimum laying out the software
+necessary to run it (e.g. versions of python and numpy, or a Makefile and
+compiler version) should be required.  That said, we recognize that some bit rot
+is expected with academic code; once the student who wrote the code graduates
+and moves on, it's very possible that there is no one else to maintain it.
+Solving the long-term support and maintenance problem for academic software is a
+major problem, well beyond the scope of this manuscript.
+
+\section{Summary}
+
+Given the importance of software to modern science and the challenges in writing
+high quality code, virtually all working scientists will need to ask for help
+from time to time. The keys to doing so effectively are remarkably simple: ask
+your questions succinctly and only after putting in some effort to solve the
+problem yourself, be considerate of other people's time and effort, and look for
+opportunities to contribute back to the community. Developers should make it
+easy for users to behave this way by providing clear documentation (including
+the appropriate venues to ask for help), encouraging participation, and
+insisting on respectful behavior from users and developers alike.
 
@@
 
+The cover image was created by artistfymo (\url{https://linktr.ee/ArtistFrmYO}).
+
 % We suggest you preserve this comment:
@@
 Alan Grossfield serves as a consultant to two companies: Moderna, Inc.
-and Atelerix Life Sciences.
+and Atelerix Life Sciences, and is a shareholder in Atelerix Life Sciences.
 

### 2022-01-05 paper.tex
added request for plain text bug reports

@@
 
+    \item For a command line tool, provide the command line and configuration
+    information in plain text, \emph{NOT} as a screen shot. A screen shot is
+    often harder to read, and at a bare minimum forces the developer to retype
+    your command instead of copying and pasting, which makes it harder for them
+    to spot possible typos, etc.
+
 \end{itemize}

### 2022-01-17 response-letter.tex
a couple more tweaks to the response letter

@@
 somewhat embarrassed I didn't think to add them myself, since they clearly
-influenced my thinking when I read them. I would argue that the checklists make
+influenced my thinking. I would argue that the checklists make
 a good short punchy summary, as the reviewer suggested. However, I agree having
@@
 about specific software engineering technologies here, though, just because I
-feel like it's a side topic I couldn't do justice to.}
+feel like it's a topic I couldn't do justice to without doubling the length of
+the paper.}
 
@@
 
-\response{This is a good point (although I'd argue that developers ought to be a bit more courteous than that).
-I added a couple of sentences at the end of the bug report section, appended to the paragraph about feature requests.}
+\response{This is a good point (although I'd argue that developers ought to be a
+bit more courteous than that). I added a couple of sentences at the end of the
+bug report section, appended to the paragraph about feature requests.}
 

### 2022-01-17 paper.tex
added ref to Lindahl's paper

@@
 This article is written from a the perspective of a developer of academic
-open-source scientific software, based on my personal experience as a developer
+open source scientific software, based on my personal experience as a developer
 and user. I released my first piece of open source code, wham \cite{WHAM}, a
@@
 reminder that the people providing the support are likely volunteers --- are
-not.  Regardless, we view the interaction between developers and users as an
+not.  Regardless, I view the interaction between developers and users as an
 informal social contract, where each has obligations and expectations for the
@@
 Doing these things is valuable from a practical perspective, as discussed
-elsewhere \cite{Heister-2013, Laayouni-2016}; making it easy for people to
-interact with us efficiently, even if it costs us some effort up front, will
-make those interactions more pleasant. Even setting that aside, it's the right
-way to do science --- FAIR principles for software development
+elsewhere \cite{Heister-2013, Laayouni-2016, Wallqvist-2019}; making it easy for
+people to interact with us efficiently, even if it costs us some effort up
+front, will make those interactions more pleasant. Even setting that aside, it's
+the right way to do science --- FAIR principles for software development
 \cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
